### 2021-02-13 main.tex
Clean up Preliminaries
and divide introduce dedicated files for sections
Preliminaries and Architecture.

\section{Preliminaries}
% Terminology

\todo[inline]{Add: brief exposure of FOL and saturation-based proving with terms possibly emphasized.}
\todo[inline]{If space is left: define KBO.}
\todo[inline]{If space is left: show resolution and superposition rules. Explain how KBO influences them.}
\todo[inline]{Mention lcm predicate. Cite the Vampire article. Voronkov: transfinite KBO - see section Notes on implementation

Mention: Vampire sets all KBO symbol weights to 1.

In Vampire: simple TKBO - level

We use TKBO to increase the impact of predicate precedence.}
\todo[inline]{Get inspiration from PAAR paper.}

\vampire{} \cite{10.1007/978-3-642-39799-8_1}
is a representative
of the state-of-the-art \gls{fol} \gls{atp} design.
\todo{Move into Introduction.}

%\vampire{} is a saturation-based prover.
%Given a \gls{fol} problem,
%\vampire{} converts the problem into \gls{cnf}.
%and proceeds to saturate the set of clauses
%by repeated application of inference rules.

%\vampire{} searches for a proof by applying the rules
%of the superposition inference system
%\cite{10.1007/978-3-642-39799-8_1}.

% AWR candidate
Proof search in Vampire is crucially constrained
by a simplification ordering on terms.
Being refutation-based and saturation-based,
Vampire searches for a contradiction
by iteratively inferring clauses provable from the input clauses.
The inferences

explores the space of clauses provable
from the input formulas
by applying the rules of superposition inference system.
This process continues until a contradiction is inferred,
the processed clause set is saturated,
or the execution hits a resource limit.

During the proof search,
Vampire explores the space of provable clauses
by applying inference rules to clauses
that have already been proven.

% AWR candidate
Simplification ordering on terms
influences the proof search in Vampire on two levels.
First, the inferences on each clause are limited
to the selected literals.
\todo{Cite Bachmair Ganzinger or Handbook of AR, chapter 2. For literal selection. Inspiration: Selecting the selection.}
In each clause,
either a negative literal or all the maximal literals are selected.
The maximality is evaluated
according to the simplification ordering.
% Note: The selection function Total does not use the simplification ordering.
Second, the simplification ordering orients some of the equalities
to prevent superposition and equality factoring
from inferring redundant complex conclusions.
In each of these two roles,
the simplification ordering may impact the direction and,
in effect, the length of the proof search.

The inference system used in the state-of-the-art \gls{fol} \gls{atp} \vampire{}
is parameterized by a simplification ordering on terms.
\vampire{} uses the superposition inference system \cite{}.

The \gls{atp} \vampire{} uses superposition inference system.

Vampire supports two simplification term ordering classes:
\gls{lpo} and \gls{kbo}.

The superposition inference system \cite{} used in Vampire is parameterized
by symbol precedence.

% Example: It seems that a classical example for KBO an orienting equations is group theory axioms.

The choice of symbol precedence may affect the length of a proof search to a great extent.

Given a \gls{fol} problem,
determining a symbol precedence that leads to a fast proof search is a nontrivial task.

\todo[inline]{Define "CNF" as quantifier-free CNF FOL with equality.}
\todo[inline]{Define "signature".}
% https://en.wikipedia.org/wiki/First-order_logic

\todo[inline]{GCN? Backprop?, gradient descent, loss, gradient}

@@

\subsection{Learning to generate good permutations}

Generating good symbol precedences is an instance of a more general task
of generating permutations of arbitrary length.
Each input problem corresponds to an input object that specifies the length of a permutation.

Let $X$ be a set of objects.
Let $l: O \to \nat$ assign each object $x \in X$ its signature length $l(x)$.
Let $\{x_1, \ldots, x_n\}$ be the training samples.\todo{MS: je mi jasny, ze tu neni poradek, ale i tak; posledni dva odstavce jsou dost zmateny.}

\subsubsection{Binary classification}

Let $p_\theta : X \to \re$ be a binary classification model.\todo{MS: Dobry priklad na to, ktere pojmy je treba vysvetlovat. model je znamy pojem v logice, tady ale pouzivame jiny, ML vyznam. To by si zaslouzilo komentar. Navic pisem
``binary'' a ``classification'' a neni tu jasny, co je ``binary''
a co jsou ``classes''. Z pohledu ctenare, ktery o tom zatim neslysel, se misto ``binary classficiation model'' mohlo
psat ``housenka'' a vyslo by to pro nej na stejno.}
The model predicts the probability of label $1$.
Let $(x_i, y_i)$ be the training samples,
where $x_i \in X$ and $y_i \in Y = \{0, 1\}$.\todo{MS:
``training'', ``sample'' a ``loss'' jsou dalsi priklady pojmu,
ktere najednou spadly z hury.}

For a training sample $(x, y) \in \re \times \{0, 1\}$, the binary cross-entropy loss is:
$$
L(\theta) = y \log{f_\theta(x)} + (1-y) \log{(1-f_\theta(x))}
$$

If $y_i = 1$ for all $i$, then the loss can be written in a simpler form:
$$
L(\theta) = \log{f_\theta(x)}
$$

\subsubsection{Permutations}

Let $O$ be a set of objects (or contexts).
Given an object $o \in O$, let $\Sigma(o)$ be its signature.
Let $x_i = (o_i, \pi_i, \rho_i)$, where $o_i \in O$
and $\pi_i, \rho_i \in \Perm(\Sigma(o_i))$.
Let $y_i \in \{0, 1\}$.

Let $p_\theta: X \to Y$.

Loss:
$$
L(\theta) = \log{p_\theta(x)}
$$

Let $c_\theta(o, s)$ be the predicted cost of symbol $s$.
Then:
$$
p_\theta(o, \pi, \rho) = C_\theta(o, \pi) - C_\theta(o, \rho)
$$

$$
C_\theta(o, \pi) = \sum_{i=1}^{l(o)}{c(o, \pi_i)}
$$

\begin{align*}
L(\theta) &= \log{p_\theta(o, \pi, \rho)} \\
&= \log{(C_\theta(o, \pi) - C_\theta(o, \rho))} \\
&= \log{(\sum_{i=1}^{l(o)}{c(o, \pi_i)} - \sum_{i=1}^{l(o)}{c(o, \rho_i)})} \\
&= \log{\sum_{i=1}^{l(o)}{c(o, \pi_i) - c(o, \rho_i)}} \\
\end{align*}

If $c$ is differentiable,
we can propagate loss gradients into it.

\subsection{Notation}

We denote the set of all finite vectors over $\re$ by $\re^* = \bigcup_{n=1}^\infty{\re^n}$.

Dot product: $\DotProd{x}{y} = \sum_{i=1}^n x_i \cdot y_i$ for $x, y \in \re^n$.

For any $n \in \nat$, we denote the set of all permutations over the set $\{1, \ldots, n\}$ by $\Perm{n}$.
Permuting a vector $x = (x_1, \ldots, x_n) \in \re^n$ by a permutation $\pi = (\pi_1, \ldots, \pi_n) \in \Perm{n}$ yields the vector $\pi(x) = (x_{\pi_1}, \ldots, x_{\pi_n})$.
$\inv{\pi}$ denotes the permutation inverse to $\pi$.
$\inv{\pi} = ...$.
\todo{Define permutation inversion.}
\todo{MS: Mozna ani ne. Ono je obcas tezky zavest a pouzivat presnout matematickou notaci, ktera se navic muze stavat neprehlednou, kvuli te presnosti. Na druhou stranu, intuitivne, kazdy vi, co by to mela byt inverzni permutace. Takze nerikam psat nepresne veci, ale obcas jde
zustak vagni s tim, ze vim, ze existuje zpusob jak to udelat presnym.
A povrchni ctenar to nemusi resit a matematicky zdatny by si ten detail odvodil spravne ...}

\newcommand{\Prec}{\pi}
\newcommand{\PrecBetter}{\pi}
\newcommand{\PrecWorse}{\rho}

\subsection{Overview}

Let $\cnf$ be the set of all \gls{cnf} problems.
Let $P \in \cnf$ be an arbitrary problem.
\todo{Consider using $F$ (as formula) instead of $P$ to free up P for probability.}
Let $\signature{P}$ be the signature of $P$, that is a list of all non-logical symbols (predicates and functions)
\todo{Should we rather use either predicates or functions? Or use a compound signature?}
that appear in the problem.
Let $n = \card{\signature{P}}$.
A symbol precedence $\Prec \in \Perm{n}$ for problem $P$ is a permutation of the symbols of $P$.

\newcommand{\Solver}{S}
\newcommand{\SolverRun}[2]{\Solver(#1, #2)}

Let $\Solver$ be an \gls{atp} with a fixed computation environment, configuration and time limit.
The computation of $\Solver$ is assumed to be parameterized by symbol precedence.
\todo{Discuss. What does this mean?}
$\SolverRun{P}{\Prec}$ denotes the result of a proof attempt on problem $P$ with symbol precedence $\Prec$.
$\SolverRun{P}{\Prec} = \top$ if $S$ solves $P$ within the time limit, and $\SolverRun{P}{\Prec} = \bot$ otherwise.
\todo{MS: vadi zminit pocet iteraci smycky tak, jako v minulem paperu?}
Since the result is generally non-deterministic in case the time limit is specified in wall clock units,
we consider $\SolverRun{P}{\Prec}$ to be a random variable with a Bernoulli distribution.
\todo{Cite?}
We denote the probability of a successful proof search $\Prob{\SolverRun{P}{\Prec} = \top}$.
\todo{MS: asi bych se primlouval za to v teto fazi o Bernoullim pomlcet.
Otazky jsme generovali tak, ze proste porovnavame pocty iteraci saturatcni smycky a pocitame s tim, ze beh je deterministicky.
To, ze se pozdeji, pri evalu, pouzije perspektiva nahodnosti
bych uz chapal jako metodiku experimentu a teorii tim nekomplikoval.}

The main task this paper deals with is this:
Given a problem $P \in \cnf$, generate a symbol precedence $\Prec \in \Perm{n}$ that maximizes $\Prob{\SolverRun{P}{\Prec} = \top}$.
MS: (Nehlede na to, ze main task by nejspis mel byt vysvetlen mnohem drive...) Asi bych to takhle nestavel. Jak pisu vyse, precijen je
beh dokazovace deterministicky a brat to nejak Bayesovsky by sice mohlo 
byt zajimave, ale jeste mnohem tezsi na obhajeni jako ``svetonazor''.
Co proste priznat nasledujici assumption:
verime, ze pokud bude model spravne odpovidat na otazky $(P, \pi_1, \pi_2)$ (s pravdepodobnosti $> \SI{50}{\percent}$), povede podle nej vybrana permutace minimalizujici \eqref{eq:model_cost} k rychlemu vyreseni problemu (a tim i idealne k vyreseni problemu, ktere byly drive out of reach.)\todo{MS: koukni prosim, jestli nejaky takovy mame. Tj zadna random permutace ho nevyresila, ale naucena sit pak ano!}

Instead of modeling $\Prob{\SolverRun{P}{\Prec} = \top}$ directly,
the system described in this paper predicts, given a problem $P$ and a pair of precedences $\PrecBetter, \PrecWorse \in \Perm{n}$,
the probability\todo{MS: tohle je jeden z oficialnich zpusobu,
jak vysvetlovat neuronky, ale neni nijak jasne do jake miry byva
$\sigma(\mathit{logits})$ blizko nejake pravdepodobnosti. Za mne by klidne stacilo mluvit o tom, ze se snazime klasifikovat co nejvic dvojic
spravne a ze na to pouzivame standard techologii neuronek (Binary Cross Entropy loss (against the final layer’s sigmoid non-linearity)). Tj opet hlasuji za to nemluvit zde o pravdepodobnosti. :)}
\todo{MS: jo, ale pak bude nekde dobre zminit, ze jsme si vedomi, ze dvojice skoro jiste neobsahuji stejne mnozstvi ``signalu'', ze ktereho je mozne se ucit. Je to proste vlastnost vybrane metody a v prumeru to nevadi.}
that $\PrecBetter$ outperforms $\PrecWorse$ in terms of success of proof search. This proxy task allows learning from pairs of precedences that both yield a successful proof search but differ in the number of steps taken in the proof search.
\todo{Explain more.}

% By taking the pairs, we get more data esp. from easy problems.
% Consider weighting the samples by the result type: quantitative / qualitative etc.

% TODO: Experiments: Increase depth and message size

\begin{figure}[h]
\caption{System overview}
\label{fig:SystemOverview}
\centering
\usetikzlibrary{shapes}
\tikzstyle{object} = [rectangle, draw]
\tikzstyle{input} = [ellipse, draw]
\tikzstyle{output} = [ellipse, draw]

\begin{tikzpicture}[node distance = 0.5 and 1, ->]
% https://tex.stackexchange.com/a/332796/202639

\node (problem) [input] {\gls{cnf} problem $P$};
\node (symbol embeddings) [object] [below=of problem] {Symbol embeddings};
\node (symbol costs) [object] [below=of symbol embeddings] {Symbol costs};
\node (symbol precedence) [output] [below=of symbol costs] {Symbol precedence};

\draw (problem) to node [left] {GCN} (symbol embeddings);
\draw (symbol embeddings) to node [left] {MLP} (symbol costs);
\draw (symbol costs) to node [left] {Order symbols by their costs} (symbol precedence);

\node (PrecBetter) [input] [right=of problem] {Precedence $\PrecBetter$};
\node (PrecBetterInv) [object] [below=of PrecBetter] {$\inv{\PrecBetter}$};
\draw (PrecBetter) to node [right] {Invert} (PrecBetterInv);

\node (PrecWorse) [input] [right=of PrecBetter] {Precedence $\PrecWorse$};
\node (PrecWorseInv) [object] [below=of PrecWorse] {$\inv{\PrecWorse}$};
\draw (PrecWorse) to node [right] {Invert} (PrecWorseInv);

\node (PrecDiff) [object] [right=2 of symbol costs] {$\inv{\PrecWorse} - \inv{\PrecBetter}$};
\node (PrecDiffNormalized) [object] [below=of PrecDiff] {Normalized};
\node (PrecPairCost) [object] [below=of PrecDiffNormalized] {Precedence pair cost};
\node (loss) [output] [below=of PrecPairCost] {Loss};

\draw (PrecBetterInv) to (PrecDiff);
\draw (PrecWorseInv) to (PrecDiff);
\draw (PrecDiff) to node [right] {Normalize} (PrecDiffNormalized);
\draw (PrecDiffNormalized) to (PrecPairCost);
\draw (PrecPairCost) to (loss);

\draw (symbol costs) to (PrecPairCost);

\end{tikzpicture}
\end{figure}

Fig.~\ref{fig:SystemOverview} \todo{MS: Na figures/tables bez referneci v textu se pohlizi obzvlast nelibe. Predpokladam, zes planoval pozdeji popsat, presto zmninuji. Konkretne k Fig.~\ref{fig:SystemOverview}:
Napada me: neslo by precedenci definovat tak, aby uz byla tim $\pi^{-1}$? Invertovani zabira v obrazku skoustu mista, ale pritom
jde o technicky detail a hlavni myslenku nijak neosvetluje...}

\subsection{Symbol cost model}

The symbol costs are modeled by a \gls{rgcn}.\cite{Schlichtkrull2017}
\todo{Can I use the term \gls{rgcn} when I use a modified version? Namely, the activation is called earlier.}
The model is a stack of graph convolutional layers.
Each layer consists of one differentiable module for each edge type.
Each module is a dense layer.
$$
h_i^{(l+1)} =
\mathrm{LayerNorm} \Parentheses{h_i^{(l)} + \sum_{r \in \mathcal{R}} \sigma \Parentheses{\sum_{j \in \mathcal{N}_i^r} \frac{1}{c_{ji}} h_j^{(l)} W_r^{(l)}}}
$$
\todo{Include dropout?}
% Inspiration: https://ufal.mff.cuni.cz/~straka/courses/npfl114/1920/slides.pdf/npfl114-07.pdf - slide 27 - Transformer
$h_i^{(l)}$ denotes the embedding of node $i$ at layer $l$.
$\mathcal{R}$ denotes the set of all relations in the graph.
$\sigma$ denotes the activation function, for example the \gls{relu}.
$\mathcal{N}_i^r$ denotes the set of neighbors of node $i$ under relation $r$.
$c_{j, i, r} = \sqrt{\card{\mathcal{N}_j^r}} \sqrt{\card{\mathcal{N}_i^r}}$ is a normalization constant.\cite{kipf2017semisupervised}
$W_r^{(l)}$ is a trainable weight matrix representing relation $r$ at layer $l$.

Residual...
Layer norm...

\todo{MS: neni uplne jiste, ze ``CADE people'' oceni presny zapis 
strutury site, vcetne prvku jako $\mathrm{LayerNorm}$ ve vzorecku.
Nerikam, ze bychom to meli uplne zatajit, ale popis, ktery je zde, spis
patri do appendixu (ML paperu). Na te siti vlastne nic originalniho neni,
tak neni treba vsechno vysvetlit do detailu. Zajimavejsi tu
potom je az to, jak se problem $P$ ``otiskne'' v topologii grafu. 
To uz originalni je (a soutezi to s representacemi jako ta Mirkova)
a popisujes to pak dal.}

\subsection{Loss function} \todo{MS: tady jsem zatim skoncil...}
\label{sec:loss}

Let $\cnf$ be the set of all \gls{cnf} problems.
Let $P \in \cnf$ be an arbitrary problem.
Let $\signature{P}$ be the signature of $P$.
Let $n = \card{\signature{P}}$.

Given a problem $P$, a symbol cost model $c : \cnf \to \re^*$
\todo{Confusing?}
returns a vector of costs of the symbols of $P$:
$$
c(P) = (c_1, \ldots, c_n)
$$

Let $\Prec \in \Perm{n}$ be a precedence of symbols of problem $P$.
Let $k(n) = \frac{2}{n(n+1)}$ be a normalization factor for precedences of length $n$.
The symbol cost model $c$ can be extended to a precedence cost model $C$
by taking the weighted sum of symbol costs:
$$
C(\Prec | P)
= k(n) \sum_{i=1}^n c_{\Prec_i} \cdot i
= k(n) \sum_{i=1}^n c_i \cdot \inv{\Prec}_i
= k(n) \DotProd{c(P)}{\inv{\Prec}}
$$
Note that the value of $C$ is minimized by a precedence that orders the symbols by their costs in non-increasing order.
\todo{Discuss, prove.}

Fixing the normalization factor $k(n)$ to the value $\frac{2}{n(n+1)}$
ensures that costs of precedences of various lengths are commensurable.
Observe namely that if $c$ assigns each symbol the same cost $c_0$,
then the precedence cost is equal to $c_0$ irrespective of the signature length $n$:
\begin{equation} \label{eq:model_cost}
C(\Prec | P) = \frac{2}{n(n+1)} \sum_{i=1}^n c_0 \cdot i = c_0
\end{equation}

Let $\PrecBetter, \PrecWorse \in \Perm{n}$ be two symbol precedences over $P$.
Let $\PrecBetter$ yield a faster proof search on $P$ than $\PrecWorse$, denoted as $\Better{\PrecBetter}{\PrecWorse}$.
The precedence cost model can be extended to predict the probability that $\PrecBetter$ is better than $\PrecWorse$:
\todo{Fix this. Ensure the polarity is sound.}
\begin{align*}
p(\Better{\PrecBetter}{\PrecWorse} | P)
&= \sigmoid \SquareBracket{C(\PrecWorse | P) - C(\PrecBetter | P)} \\
%&= \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)} \\
&= \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}}
\end{align*}
It is easy to see\todo{MS: Bez znalosti $\sigmoid$ uplne ne ;)} that $p(\Better{\PrecBetter}{\PrecWorse} | P) > 0.5$ if and only if $C(\PrecWorse | P) > C(\PrecBetter | P)$,
and that the value of $p$ can be increased by increasing $C(\PrecWorse | P)$ or decreasing $C(\PrecBetter | P)$.

We consider the difference $C(\PrecWorse | P) - C(\PrecBetter | P)$
to be the "logit"\todo{MS: Logit je taky tezky pojem, urcite ne automaticky znamy u ATP people.} score of the pair of precedences $\PrecBetter, \PrecWorse$,
and denote it by $C(\Better{\PrecBetter}{\PrecWorse} | P)$.

We define the loss of $c$ on the training sample $(\PrecBetter, \PrecWorse, P)$ as binary cross-entropy
with ground truth $\Better{\PrecBetter}{\PrecWorse}$:

\begin{align*}
L(\PrecBetter, \PrecWorse, P)
&= - \log p(\Better{\PrecBetter}{\PrecWorse} | P) \\
&= - \log \sigmoid \SquareBracket{k(n) \DotProd{c(P)}{\inv{\PrecWorse} - \inv{\PrecBetter}}} \\
&= - \log \sigmoid \SquareBracket{k(n) \sum_{i=1}^n c_i \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)}
\end{align*}

The loss is differentiable with respect to the symbol costs:
\begin{align*}
\frac{\partial L}{\partial c_i}
&= -\sigmoid(-C(\Better{\PrecBetter}{\PrecWorse} | P)) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i) \\
&= (p(\Better{\PrecBetter}{\PrecWorse} | P) - 1) \cdot k(n) \cdot (\inv{\PrecWorse}_i - \inv{\PrecBetter}_i)
\end{align*}

This means that it is possible to backpropagate the loss gradient into the symbol cost model. \todo{MS: tohle bude tezky,
ale ackoliv ML people by tohle uz asi chapali, ATP crowd spis bude
potrebovat obsirnejsi vysvetleni.}

\subsubsection{Generalization}

Note that this approach is generally applicable to all problems where it is necessary to generate permutations of arbitrary length.

\todo[inline]{Finish.}

\subsection{Symbol cost model}

\subsection{Graph representation of problems}

\newcommand{\ntype}[1]{\texttt{#1}}
\newcommand{\etype}[1]{\texttt{#1}}
\newcommand{\epos}{\etype{pos}}
\newcommand{\eneg}{\etype{neg}}

Every \gls{cnf} problem can be represented by a \gls{hin} with the network schema shown in \cref{fig:CnfSchema}.\todo{MS: Mel jsem chut to opravit na velke "Fig.", ale s makrem se mi hadat nechce ;)}
% Cite: https://www.kdd.org/exploration_files/V14-02-03-Sun.pdf
% Referenced from: https://docs.dgl.ai/en/0.5.x/generated/dgl.DGLHeteroGraph.metagraph.html

\begin{figure}[ht]
\caption{CNF network schema}
\label{fig:CnfSchema}
\centering
\tikzstyle{token} = [rectangle, draw]

\begin{tikzpicture}[node distance = 1 and 2, ->]
% https://tex.stackexchange.com/a/332796/202639

% Node and edge types:
% https://docs.google.com/spreadsheets/d/1PCPHEgk6vLxpdpcvB_PGoLx7p4DID6WtvVWy2mDuv4A/edit?usp=sharing

\node (formula) [token] {\ntype{problem}};
% TODO: Consider removing the Formula nodes.
\node (clause) [token, below=of formula] {\ntype{clause}};
\node (atom) [token, below left=of clause] {\ntype{atom}};
\node (equality) [token, below right=of clause] {\ntype{equality}};
\node (predicate) [token, left=of atom] {\ntype{predicate}};
\node (argument) [token, below=of atom] {\ntype{argument}};
\node (term) [token, below=of argument] {\ntype{term}};
\node (function) [token, left=of term] {\ntype{function}};
\node (variable) [token, right=of term] {\ntype{variable}};

\draw (formula) to node [right] {\etype{contains}} (clause);
\draw (clause) to [bend right] node [above] {\epos{}} (atom);
\draw (clause) to [bend left] node [below] {\eneg{}} (atom);
\draw (clause) to [bend left] node [above] {\epos{}} (equality);
\draw (clause) to [bend right] node [below] {\eneg{}} (equality);
\draw (clause) to node [above] {\etype{binds}} (variable);
\draw (atom) to node [above] {\etype{atom\_applies}} (predicate);
\draw (atom) to node [left] {\etype{atom\_has}} (argument);
\draw (equality) to node {\etype{equalizes}} (term);
\draw (equality) to node {\etype{equalizes}} (variable);
\draw (argument) to [bend right] node [left] {\etype{is}} (term);
\draw (argument) to [loop left] node [left] {\etype{precedes}} (argument);
\draw (argument) to node [below] {\etype{is}} (variable);
\draw (term) to node [above] {\etype{term\_applies}} (function);
\draw (term) to [bend right] node [right] {\etype{term\_has}} (argument);

\end{tikzpicture}
\end{figure}

The network of a problem is a directed graph $G = (V, E)$.
The function $\tau: V \rightarrow A$ maps the nodes to their types:

$$
A = \{\ntype{problem}, \ntype{predicate}, \ntype{function}, \ntype{clause}, \ntype{atom}, \ntype{equality}, \ntype{term}, \ntype{variable}, \ntype{argument}\}
$$

The function $\phi: E \rightarrow E$ maps the edges to their types.

\begin{itemize}
\item The graph representation of problem $P$ contains exactly one root node of type \ntype{problem}.
\item Each clause is represented by a \ntype{clause} node connected to the root \ntype{problem} node.
\item Each atom is represented by an \ntype{atom} node (in case the atom is not an equality)
or an \ntype{equality} node (in case the atom is an equality).
For each literal occurrence there is an edge connecting the respective atom to the respective clause.
The type of the edge corresponds to the polarity of the literal:
\epos{} for positive literal and \eneg{} for negative literal.
\item Each \ntype{equality} node is connected to two nodes that represent the operands,
each of which is of type \ntype{term} or \ntype{variable}.
The commutativity of the equality operator is reflected by the fact that the operand edges are not ordered.
\item Each \ntype{atom} node is connected to one \ntype{predicate} node that represents the predicate symbol being applied by this atom.
Note that these edges connect the applications of a predicate symbol across the whole problem.
\item Each \ntype{atom} node is connected to zero or more \ntype{argument} nodes that represent the argument positions of the atom.
\item Each pair of \ntype{argument} nodes that correspond to consecutive argument positions is connected by an edge.
\item Each \ntype{argument} node is connected to a node that represents the argument term,
which is either a \ntype{term} node or a \ntype{variable} node.
\item Each \ntype{term} node is connected to one \ntype{function} node that represents the function symbol being applied by this term.
\item For each variable, there is an edge connecting the \ntype{clause} node of the clause that binds the variable to the \ntype{variable} node that represents the variable.
\end{itemize}

\todo[inline]{Abstract from the edge types. Focus on the main idea. Technical details may be in technical report or appendix.}

Various levels of term sharing are possible. [continue]

\todo[inline]{Give an example of a problem and its graph.}

\subsection{Generating precedences}
\label{sec:generating}

When presented with a \gls{cnf} problem $P$ with signature $\symbols$,
the recommender produces a precedence $\pi$ of symbols from $\symbols$.
In order to generate the precedence,
the recommender first assigns a cost value to each symbol
by invoking a symbol cost model (a trained \gls{gcn}) on a graph representation of $P$.
Ordering the symbols by their predicted costs in nondecreasing order yields the precedence $\pi$.

\subsection{Training}

The training is performed using a fixed \acrlong{atp}.
The configuration of the prover,
including for example saturation algorithm, age-weight ratio, and term ordering scheme,
is arbitrary\todo{MS: tohle slovo je prilis silny. Veta by spis mela vysvetlit, ze jsme si neco vybrali, protoze je nam to v postate jedno. Ale zaroven si uvedomujeme, ze to nejak ovlivni vysledek a ze v ultimatnim super-toolu, by se na vsech ostatni paramatrech dalo conditionovat.} and fixed.
The prover needs to support refutational reasoning on \gls{cnf} problems
and use a term simplification ordering parameterized by a symbol precedence.
For example, the \glspl{atp} \vampire{} and E satisfy this requirement.
Note that both of these provers support the \gls{kbo} and \gls{lpo} term simplification ordering schemes,
and that each of these schemes is parameterized by a symbol precedence.

As outlined in \cref{sec:generating},
it is necessary to train a model of symbol costs.
The model is a \gls{gcn}.

\todo[inline]{Include a metagraph and an example graph of a problem.}
\todo[inline]{Compare to HINs}
% https://medium.com/@jason_trost/heterogeneous-information-networks-and-applications-to-cyber-security-23b245461adb

Since it is not obvious what the target symbol cost values should be,
the symbol cost model is not trained directly.
Instead, it is plugged into a classifier that predicts,
given a problem $P$ and a pair of precedences $\pi_0, \pi_1$,
which of the two precedences yields a better performance on $P$.

The symbol cost model is trained by plugging the model into a classifier
that is trained to predict which of a pair of precedences
yields a better performance on an arbitrary problem.

\begin{figure}[ht]
\caption{Architecture overview}
\centering
\digraph[scale=0.4]{precedencepairclassifierdetailed}{
	graph [splines=ortho];
	node [shape=renctangle, fontsize=20];
	edge [fontsize=20];
	fol [label="FOL problem", shape=oval];
	pi0 [shape=oval, label=<&pi;<SUB>0</SUB>>];
	pi1 [shape=oval, label=<&pi;<SUB>1</SUB>>];
	invpi0 [label=<&pi;<SUB>0</SUB><SUP>-1</SUP>>];
	invpi1 [label=<&pi;<SUB>1</SUB><SUP>-1</SUP>>];
	cnf [label="Clause normal form (CNF)"];
	symbolembeddings [label="Symbol embeddings"];
	symbolcosts [label="Symbol costs"];
	pi1pi0 [label="Inverse precedence difference"];
	normalized [label="Normalized inverse precedence difference"];
	paircost [label="Precedence pair cost"];
	fol -> cnf [xlabel=" Vampire "];
	cnf -> symbolembeddings [xlabel=< <B>Graph Convolution Network</B> >];
	symbolembeddings -> symbolcosts [xlabel=< <B>Feed-forward neural network</B> >, style=bold];
	symbolcosts -> paircost [style=bold];
	paircost -> loss [xlabel=" Binary cross-entropy ", style=bold];
	loss [label="Loss", shape=oval];
	pi0 -> invpi0 [xlabel=" Invert "];
	pi1 -> invpi1 [xlabel=" Invert "];
	invpi0 -> pi1pi0;
	invpi1 -> pi1pi0;
	pi1pi0 -> normalized [xlabel=" Normalize "];
	normalized -> paircost;
	symbolprecedence [label="Symbol precedence", style=dashed];
	symbolcosts -> symbolprecedence [xlabel=" Order symbols by their costs ", style=dashed];
}
\end{figure}

\subsection{Layers}
'
\begin{enumerate}
\item Problem -> symbol embeddings
\item Symbol embedding -> symbol cost
\item Symbol costs -> precedence cost
\end{enumerate}

\subsection{Cost models}

%Let $\CostSym: \symbols \rightarrow \re$ be a differentiable symbol cost model.

We define the precedence cost:
$$
\CostPrec(\pi) =
C \sum_{1 \leq i \leq n} \CostSym(\pi(i)) \cdot i =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot \inv{\pi}(s_i)
$$
Precedence cost is minimized by $\pi$ that orders the symbols by their costs in non-increasing order
($\forall (1 \leq i < j \leq n) . (\CostSym(\pi(i)) \geq \CostSym(\pi(j)))$).

Note that we can weight the symbols with an arbitrary nondecreasing function $f$ of symbol index:
$$
\CostPrec(\pi) =
C \sum_{1 \leq i \leq n} \CostSym(\pi(i)) \cdot f(i) =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot f(\inv{\pi}(s_i))
$$

We set $C = \frac{2}{n(n+1)}$ so that $\CostSym(s) = 1$ for all $s$ implies $\CostPrec(\pi) = 1$ for all $\pi$.

% Note that we use this orientation because the TensorFlow metric BinaryCrossentropy classifies 0 as negative and we use the value 0 for "failed to classify" logits.
Given a pair of precedences $\pi_0, \pi_1$,
we define the log-odds of the event "$\pi_0$ is better than $\pi_1$":
$$
\CostPrecPair(\pi_0, \pi_1) =
\CostPrec(\pi_1) - \CostPrec(\pi_0) =
C \sum_{1 \leq i \leq n} \CostSym(s_i) \cdot [\inv{\pi_1}(s_i) - \inv{\pi_0}(s_i)]
$$
Clearly $\CostPrecPair(\pi_0, \pi_1) > 0$ iff $\CostPrec(\pi_0) < \CostPrec(\pi_1)$.
For a pair of precedences about which we know that $\pi_0$ is better than $\pi_1$,
we want $\CostPrecPair(\pi_0, \pi_1) > 0$.

We model the probability of the event "$\pi_0$ is better than $\pi_1$"
by the sigmoid of $\CostPrecPair(\pi_0, \pi_1)$:
$$
p(\pi_0, \pi_1) = \sigmoid(\CostPrecPair(\pi_0, \pi_1))
$$

We use the binary cross-entropy loss to train the model.
Given a pair of precedences such that $\pi_0$ is better than $\pi_1$,
the loss is as follows:
$$
Loss(\pi_0, \pi_1) = -\log(\sigmoid(\CostPrecPair(\pi_0, \pi_1)))
$$

>>>>

\section{Preliminaries}

\input{preliminaries}

@@

\input{architecture}
